---
title: "HW 1"
output: github_document
date: "2026-09-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Calling necessary packages.
```{r, message=FALSE}
library("tidyverse") #for ggplot
library("palmerpenguins")
```

### Problem 1

Load the penguins dataset.
```{r, results= "hide"}
data("penguins", package = "palmerpenguins")

str(penguins)
```

In the `penguins` dataset, there are `r nrow(penguins)` rows and `r ncol(penguins)` columns of data. One row corresponds to one penguin. The variables are `r colnames(penguins)`. The `species` column lists one of three penguin species for each individual penguin: Adelie, Chinstrap, and Gintoo. The `island` columns contains the island of each penguin: Biscoe, Dream, and Torgersen. `bill_length_mm` contains each penguin's bill length in mm `bill_depth_mm` states the bill depth in mm. Both bill length and bill depth are expressed with decimals. `flipper_length_mm` and `body_mass_g` both express numeric values that measure the length of a penguin's flipper, and the body mass of each penguin respectively. `sex` is a factor with two levels: "female" and "male". The `year` column expresses the year.The mean of `flipper_length_mm` is `r mean(penguins$flipper_length_mm, na.rm = TRUE )` mm.

Scatterplot of `flipper_length_mm` and `bill_length_mm`
```{r}
penguin_scatter = ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color = species)) + 
  geom_point(na.rm = TRUE)

penguin_scatter
```
 
Save plot. 
```{r}
ggsave("penguin_scatter.pdf", height = 4, width = 6)
```