---
title: "HW 1"
output: github_document
date: "2026-09-24"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Calling necessary packages.
```{r, message = FALSE}
library("tidyverse") #for ggplot
library("palmerpenguins")
```

### Problem 1

Load the penguins dataset.
```{r, results = "hide", message = FALSE}
data("penguins", package = "palmerpenguins")

str(penguins)
```

In the `penguins` dataset, there are `r nrow(penguins)` rows and `r ncol(penguins)` columns of data. One row corresponds to one penguin. The variables are `r colnames(penguins)`. The `species` column lists one of three penguin species for each individual penguin: Adelie, Chinstrap, and Gintoo. The `island` columns contains the island of each penguin: Biscoe, Dream, and Torgersen. `bill_length_mm` contains each penguin's bill length in mm `bill_depth_mm` states the bill depth in mm. Both bill length and bill depth are expressed with decimals. `flipper_length_mm` and `body_mass_g` both express numeric values that measure the length of a penguin's flipper, and the body mass of each penguin respectively. `sex` is a factor with two levels: "female" and "male". The `year` column expresses the year.The mean of `flipper_length_mm` is `r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)` mm.

Scatterplot of `flipper_length_mm` and `bill_length_mm`
```{r}
penguin_scatter = ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color = species)) + 
  geom_point(na.rm = TRUE)

penguin_scatter
```
 
Save plot. 
```{r}
ggsave("penguin_scatter.pdf", height = 4, width = 6)
```

### Problem 2

Creating a new data frame. 
```{r}
set.seed(123)

df_2 = tibble(
  vec_random = rnorm(10),
  vec_logic = vec_random > 0,
  vec_char = c("a","b","c","d","e","f","g","h","i","j"),
  vec_fact = factor(c("easy","easy","medium","hard","easy","hard","medium","hard","medium","easy"))
)

print(df_2)
```
Taking the mean of each variable
```{r}
mean(pull(df_2, vec_random))
mean(pull(df_2, vec_logic))
mean(pull(df_2, vec_char))
mean(pull(df_2, vec_fact))
```
* Taking the mean only works for `vec_random` and `vec_logic`. `vec_random` produces numerical values that can be used to take the mean. In logical vectors, TRUE and FALSE can be interpreted as 1 for TRUE and 0 for FALSE, thus giving us the ability to take the mean.

`as.numeric` to our logical, character, and factor variables.
```{r, results = "hide", warning = FALSE}
as.numeric(pull(df_2, vec_logic))
as.numeric(pull(df_2, vec_char))
as.numeric(pull(df_2, vec_fact))
```
When you convert logic variables to numeric ones, we receive 1's and 0's since TRUE and FALSE logical values care treated as 1 and 0. For factor variables, the strings inside the vector are assigned integer levels. In contrast, character variables have no underlying mathematical meaning.Thus, we can take the means of logical and factor variables, but not character variables.